---
title: 'Geog6300: Final Project: Modelling Water Surface Salinity from Remote Sensing based Reflectance Data'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Full assignment available at https://docs.google.com/document/d/1-e6G9iR-WLWqQQpj3vls4-BNGaeSFPv4OalxWQoX52M/edit?usp=sharing

# install.packages("ggpubr")

#Importing Libraries
library(tidyverse)
library(pander)
library(ggpubr)
library(corrplot)
library(car)
library(lmtest)

# loading data
salinity_data <- read.csv('s2boa_salinity.csv')

# check for na values (already removed while compiling dataset, double checking here) and drop tuples with na values
is.na(salinity_data)
salinity_data <- na.omit(salinity_data)
```

Your name: Chintan B. Maniyar

## Research question: 
* Water Surface Salinity (WSS) is a measure of the amount of dissolved oxygen in the water and has no evident color signal. This project aims to answer the following two  questions:  
1. Is it possible to estimate water surface salinity by remote sensing methods using satellite data, specifically along the coast of Georgia? If yes, what wavelength is sensitive to WSS variation? 

2. Moreover, since WSS depends highly on the turbidity and suspended particles in the water, can remote sensing based indices/quantities such as Color Dissolved Organic Matter (CDOM), Normalized Difference Chlorophyll Index (NDCI) and Normalized Suspended Material Index (NSMI) be used as proxy to estimate WSS from satellite data? 


## Data
* The dataset for this assignment is a combination of in-situ and coincident satellite data. In-situ WSS and water surfce temperature values were recorded for the Georgian coast by field sampling as well as Research Vehicle Savannah. Co-incident Sentinel-2 images captured within a 30 minute window of in-situ sampling were acquired for 2018-2020 period. Pixel values for all 13 bands were extracted from these images based on the ground sampling locations. Hence, this dataset looks something like this: for each sampling point: Surface Reflectance (SR) values from Sentinel-2 and corresponding WSS value recorded on ground. 
* From the raw bands, the aforementioned indices (CDOM, NDCI, NSMI) are also calculated
* Additional band ratios are calculated based on literature survey
* Source of Ground Data: Georgia Coastal Ecosystems (GCE-LTER)
* Satellite data was curated by downloading coincident freely available Sentinel-2 Imagery

* Following chunk shows the sample of the dataset:

```{r}
pander(head(salinity_data))
```

## Methods
* As discussed thus far, the main goal of this project is to investigate if it is possible to estimate WSS entirely using satellite data. This essentially means investigating the relationship and underlying pattern (if any) between the satellite recorded SR values and the ground sampled WSS values. 

* To answer question 1, I propose to use a correlation matrix with the variables: Bands 1 through 12, all indices and band ratios and ground sampled WSS values. I choose correlation since all variables are continuous. Based on whether the data is parametric or non-parametric, I would choose pearson or spearman correlation respectively. 
This would:    
  1. Show any underlying correlation between any of the bands and salinity values, hence answering the question if a particluar wavelength in the optical spectrum is sensitive to WSS variation. 
  2. Also show underlying correlation between WSS and all indices and band ratios, enabling me to understand better if the literature reviews are relavant for my study site as well

* To answer question 2, I propose to build a multi-variate linear regression  (MLR) model with bands 1 through 12, band ratios and indices as independent variables (initially) and WSS as the dependent variable. The set of independent variables will be consequently modified by checking for multicollinearity (VIF test) and homoscedastity (bptest), in order to ensure that each independent variable is carrying unique information to the model. Finally an MLR model with a refined set of independent variables will be suggested and inferred, leading to a discussion of how far along can one get in estimating WSS values using satellite data.  

## Analysis
This section is structured into 3 subsections. While after every code chunk there will be inferential statements, the end of each subsection will summarize the purpose and overall inference of that subsection. 

### 0. Data Pre-Analysis
Before we actively start to probe our research questions, let us understand the nature of our dependent and independent variables -- whether or not they are normal. We would plot qq-plots for all of them to test their normality. If the qq-plot aligns in a straight line, it means that variable is distributed normally. If the qq-plot resembles a smile or a frown, it means the the variable is positively skewed or negatively skewed respectively. 

```{r}
b1 <- ggplot(salinity_data, aes(sample=ï..B1)) + stat_qq() + stat_qq_line() + ggtitle("Band-1")
b2 <- ggplot(salinity_data, aes(sample=B2)) + stat_qq() + stat_qq_line() + ggtitle("Band-2")
b3 <- ggplot(salinity_data, aes(sample=B3)) + stat_qq() + stat_qq_line() + ggtitle("Band-3")
b4 <- ggplot(salinity_data, aes(sample=B4)) + stat_qq() + stat_qq_line() + ggtitle("Band-4")
b5 <- ggplot(salinity_data, aes(sample=B5)) + stat_qq() + stat_qq_line() + ggtitle("Band-5")
b6 <- ggplot(salinity_data, aes(sample=B6)) + stat_qq() + stat_qq_line() + ggtitle("Band-6")
b7 <- ggplot(salinity_data, aes(sample=B7)) + stat_qq() + stat_qq_line() + ggtitle("Band-7")
b8 <- ggplot(salinity_data, aes(sample=B8)) + stat_qq() + stat_qq_line() + ggtitle("Band-8")
b8a <- ggplot(salinity_data, aes(sample=B8A)) + stat_qq() + stat_qq_line() + ggtitle("Band-8A")
b9 <- ggplot(salinity_data, aes(sample=B9)) + stat_qq() + stat_qq_line() + ggtitle("Band-9")
b11 <- ggplot(salinity_data, aes(sample=B11)) + stat_qq() + stat_qq_line() + ggtitle("Band-11")
b12 <- ggplot(salinity_data, aes(sample=B12)) + stat_qq() + stat_qq_line() + ggtitle("Band-12")
ndvi <- ggplot(salinity_data, aes(sample=NDVI)) + stat_qq() + stat_qq_line() + ggtitle("NDVI")
cdom <- ggplot(salinity_data, aes(sample=CDOM.Proxy..B3.B4.)) + stat_qq() + stat_qq_line() + ggtitle("CDOM")
ndci <- ggplot(salinity_data, aes(sample=NDCI)) + stat_qq() + stat_qq_line() + ggtitle("NDCI")
nsmi <- ggplot(salinity_data, aes(sample=NSMI)) + stat_qq() + stat_qq_line() + ggtitle("NSMI")
b43 <- ggplot(salinity_data, aes(sample=B4.B3)) + stat_qq() + stat_qq_line() + ggtitle("B4/B3")
b42 <- ggplot(salinity_data, aes(sample=B4.B2)) + stat_qq() + stat_qq_line() + ggtitle("B4/B2")
b41 <- ggplot(salinity_data, aes(sample=B4.B1)) + stat_qq() + stat_qq_line() + ggtitle("B4/B1")
salinity = ggplot(salinity_data, aes(sample=GCE_Salinity)) + stat_qq() + stat_qq_line() + ggtitle("Salinity")

ggarrange(b1, b2, b3, b4, b5, b6, nrow = 2, ncol = 3)
ggarrange(b7, b8, b8a, b9, b11, b12, nrow = 2, ncol = 3)
ggarrange(ndvi, cdom, ndci, nsmi, b43, b42, nrow = 2, ncol = 3)
ggarrange(b41, salinity, nrow = 1, ncol = 2)
# b6, b7, b8, b8a, b9, b11, b12, ndvi, cdom, ndci, nsmi, b43, b42, b41,
```

We can observe that most of the variables are not normally distributed (with exceptions such as NSMI). Moreover. our dependent variable, WSS, has an interesting qq-plot. It is on either sides of the normal line, which probably means that the variable is bimodal, likely having high values near 0 and around 25-30. Let's check that by plotting a histogram for WSS.

```{r}
ggplot(salinity_data, aes(x=GCE_Salinity)) + geom_histogram() + ggtitle("Salinity")
```

Clearly, WSS is a skewed variable. To make it a little less skewed, we will filter out salinity values greater than 2. We will discard salinity values from 0-2, as that is freshwater and we want to explore the salinity dynamics along the coastal region while focusing on the transition zone between freshwater and oceanic water. [Reference](https://salinometry.com/pss-78/)
Once filtered, we will once again check the normality of all variables before we do further analysis.

```{r}
salinity_data_filtered <- salinity_data %>%
  filter(GCE_Salinity > 2)

ggplot(salinity_data_filtered, aes(x=GCE_Salinity)) + geom_histogram() + ggtitle("Salinity")
```

While still not normal, at least the WSS variable is tending towards a unimodal distribution now. Let's check if this changed the normality of other variables.

```{r}
b1 <- ggplot(salinity_data_filtered, aes(sample=ï..B1)) + stat_qq() + stat_qq_line() + ggtitle("Band-1")
b2 <- ggplot(salinity_data_filtered, aes(sample=B2)) + stat_qq() + stat_qq_line() + ggtitle("Band-2")
b3 <- ggplot(salinity_data_filtered, aes(sample=B3)) + stat_qq() + stat_qq_line() + ggtitle("Band-3")
b4 <- ggplot(salinity_data_filtered, aes(sample=B4)) + stat_qq() + stat_qq_line() + ggtitle("Band-4")
b5 <- ggplot(salinity_data_filtered, aes(sample=B5)) + stat_qq() + stat_qq_line() + ggtitle("Band-5")
b6 <- ggplot(salinity_data_filtered, aes(sample=B6)) + stat_qq() + stat_qq_line() + ggtitle("Band-6")
b7 <- ggplot(salinity_data_filtered, aes(sample=B7)) + stat_qq() + stat_qq_line() + ggtitle("Band-7")
b8 <- ggplot(salinity_data_filtered, aes(sample=B8)) + stat_qq() + stat_qq_line() + ggtitle("Band-8")
b8a <- ggplot(salinity_data_filtered, aes(sample=B8A)) + stat_qq() + stat_qq_line() + ggtitle("Band-8A")
b9 <- ggplot(salinity_data_filtered, aes(sample=B9)) + stat_qq() + stat_qq_line() + ggtitle("Band-9")
b11 <- ggplot(salinity_data_filtered, aes(sample=B11)) + stat_qq() + stat_qq_line() + ggtitle("Band-11")
b12 <- ggplot(salinity_data_filtered, aes(sample=B12)) + stat_qq() + stat_qq_line() + ggtitle("Band-12")
ndvi <- ggplot(salinity_data_filtered, aes(sample=NDVI)) + stat_qq() + stat_qq_line() + ggtitle("NDVI")
cdom <- ggplot(salinity_data_filtered, aes(sample=CDOM.Proxy..B3.B4.)) + stat_qq() + stat_qq_line() + ggtitle("CDOM")
ndci <- ggplot(salinity_data_filtered, aes(sample=NDCI)) + stat_qq() + stat_qq_line() + ggtitle("NDCI")
nsmi <- ggplot(salinity_data_filtered, aes(sample=NSMI)) + stat_qq() + stat_qq_line() + ggtitle("NSMI")
b43 <- ggplot(salinity_data_filtered, aes(sample=B4.B3)) + stat_qq() + stat_qq_line() + ggtitle("B4/B3")
b42 <- ggplot(salinity_data_filtered, aes(sample=B4.B2)) + stat_qq() + stat_qq_line() + ggtitle("B4/B2")
b41 <- ggplot(salinity_data_filtered, aes(sample=B4.B1)) + stat_qq() + stat_qq_line() + ggtitle("B4/B1")
salinity = ggplot(salinity_data_filtered, aes(sample=GCE_Salinity)) + stat_qq() + stat_qq_line() + ggtitle("Salinity")

ggarrange(b1, b2, b3, b4, b5, b6, nrow = 2, ncol = 3)
ggarrange(b7, b8, b8a, b9, b11, b12, nrow = 2, ncol = 3)
ggarrange(ndvi, cdom, ndci, nsmi, b43, b42, nrow = 2, ncol = 3)
ggarrange(b41, salinity, nrow = 1, ncol = 2)
```

We can now see that the WSS qq-plot does not go on either sides of the normal line. All short-wavelength bands are almost unchanged. However, I feel that CDOM and band-ratios also normalized a little more than earlier, which makes me wonder if they are highly correlated to WSS. Let's plot scatter plots of each variable against WSS to explore this.

```{r}
b1 <- ggplot(salinity_data_filtered, aes(x=ï..B1, y=GCE_Salinity)) + geom_point() + ggtitle("Band-1")
b2 <- ggplot(salinity_data_filtered, aes(x=B2, y=GCE_Salinity)) + geom_point() + ggtitle("Band-2")
b3 <- ggplot(salinity_data_filtered, aes(x=B3, y=GCE_Salinity)) + geom_point() + ggtitle("Band-3")
b4 <- ggplot(salinity_data_filtered, aes(x=B4, y=GCE_Salinity)) + geom_point() + ggtitle("Band-4")
b5 <- ggplot(salinity_data_filtered, aes(x=B5, y=GCE_Salinity)) + geom_point() + ggtitle("Band-5")
b6 <- ggplot(salinity_data_filtered, aes(x=B6, y=GCE_Salinity)) + geom_point() + ggtitle("Band-6")
b7 <- ggplot(salinity_data_filtered, aes(x=B7, y=GCE_Salinity)) + geom_point() + ggtitle("Band-7")
b8 <- ggplot(salinity_data_filtered, aes(x=B8, y=GCE_Salinity)) + geom_point() + ggtitle("Band-8")
b8a <- ggplot(salinity_data_filtered, aes(x=B8A, y=GCE_Salinity)) + geom_point() + ggtitle("Band-8A")
b9 <- ggplot(salinity_data_filtered, aes(x=B9, y=GCE_Salinity)) + geom_point() + ggtitle("Band-9")
b11 <- ggplot(salinity_data_filtered, aes(x=B11, y=GCE_Salinity)) + geom_point() + ggtitle("Band-10")
b12 <- ggplot(salinity_data_filtered, aes(x=B12, y=GCE_Salinity)) + geom_point() + ggtitle("Band-11")
ndvi <- ggplot(salinity_data_filtered, aes(x=NDVI, y=GCE_Salinity)) + geom_point() + ggtitle("Band-12")
cdom <- ggplot(salinity_data_filtered, aes(x=CDOM.Proxy..B3.B4., y=GCE_Salinity)) + geom_point() + ggtitle("Band-CDOM")
ndci <- ggplot(salinity_data_filtered, aes(x=NDCI, y=GCE_Salinity)) + geom_point() + ggtitle("NDCI")
nsmi <- ggplot(salinity_data_filtered, aes(x=NSMI, y=GCE_Salinity)) + geom_point() + ggtitle("NSMI")
b43 <- ggplot(salinity_data_filtered, aes(x=B4.B3, y=GCE_Salinity)) + geom_point() + ggtitle("B4/B3")
b42 <- ggplot(salinity_data_filtered, aes(x=B4.B2, y=GCE_Salinity)) + geom_point() + ggtitle("B4/B2")
b41 <- ggplot(salinity_data_filtered, aes(x=B4.B1, y=GCE_Salinity)) + geom_point() + ggtitle("B4/B1")

ggarrange(b1, b2, b3, b4, nrow = 2, ncol = 2)
ggarrange(b5, b6, b7, b8, nrow = 2, ncol = 2)
ggarrange(b8a, b9, b11, b12, nrow = 2, ncol = 2)
ggarrange(ndvi, cdom, ndci, nsmi, nrow = 2, ncol = 2)
ggarrange(b43, b42, b41, nrow = 2, ncol = 2)
```

We can observe that Band-1 does show some positive correlation with WSS, but as we start to go towards higher wavelength, the scatter plot starts to spread out (bands 4,5) indicating decrease in correlation. Band 6 onwards it almost seems as if the scatter plot is almost all over the place, indicating very less or no correlation, with the exception of weak negative correlation for Band 12. On the other hand, the indices and band ratios, as compared to the individual bands, definitely show a pattern in the scatter plot indicating the presence of some correlation with WSS (all ecept CDOM show negatively correlation), which infact aligns well with my literature review.

> Section Summary: In this section we try to understand the nature of our data. We find out that WSS is bimodal, which makes it unsuitable to develop a model on right away, as those outliers will defintely create a bias. So we get rid of WSS values less than 2 psu and discard freshwater samples, with the motivation of looking into coastal ecosystem. We then analyze qq-plots which tell us that most of the variables are not normal and we further explore the relationship of all variables with WSS with the help of scatter plots. 


### 1. What wavelength in the optical spectrum is sensitive to WSS variation?
To figure out what region of the optical spectrum is ideal to use for estimating WSS, we will plot a correlation matrix of all bands and WSS variable. Since we saw in the previous section that most of the data are not normal, we will use Spearman Correlation.

```{r}
# filtering out only bands
salinity_bands <- salinity_data_filtered %>% 
  select(-NDVI, -CDOM.Proxy..B3.B4., -NDCI, -NSMI, -B4.B3, -B4.B2, -B4.B1, -GCE_Temperature.C.)

# confirming all we have is bands and WSS
pander(head(salinity_bands))

# now we will plot a correaltion matrix
CM <- cor(salinity_bands, method = "spearman")
corrplot(CM, method = "color", addCoef.col = "grey", number.cex = 0.7)
```

To see which bands are sensitive to WSS variation, we simply need to look at the last row (GCE_Salinity). To look at this in another interesting way, we can plot a line chart of only the last row and look for evident peaks or bumps - this would show high correlation in either direction and we can conclude that those bands, if not completely, are more sensitive to WSS variation from among all 12 bands. We will plot this line chart in the next step. 

Before that, there is also a lot of other information in this correlation plot:

* WSS variation is positively correlated with shortwavelength bands,but it starts to form an inverse relationship as we go towards high wavelength bands.
* Bands 2,3 are highly correlated 
* Bands 4,5 are highly correlated 
* Bands 6,7,8,8A are highly correlated 
* Bands 11,12 are highly correlated 

We can use this information when we build a model to predict salinity using bands as covariables. For instance, from a group of highly correlated bands, we should choose only one of them as a covariable, maybe the one which has highlest correlation with WSS. This will make sure that the model does not get confused because each covariable will then carry uniue information. 

```{r}
corr_data <- data.frame(as.data.frame(CM)$GCE_Salinity)
corr_data$Bands <- c('B1','B2','B3','B4','B5','B6','B7','B8','B8A','B9','B11','B12', 'WSS')
colnames(corr_data) <- c('Correlation', 'Bands')

# reomving last row (we don't need WSS)
corr_data <- head(corr_data, -1)

# adding a dummy column to reorder data -- this is to avoid the automatic sorting of x-axis by ggplot
corr_data$order <- c(1: length(corr_data$Bands))

ggplot(data = corr_data, aes(x=reorder(Bands, order), y = Correlation, yend=0, group = 1)) +
  xlab("Bands") +
  geom_line(color = "red") + 
  geom_point() +
  geom_hline(yintercept = 0)
  
```

Based on the graph above, which shows the variation of correlation with WSS among all bands, I would select the bands farthest from 0 in either directions as the ones sensitive to WSS variation. Following that criteria, we can select Bands 1, 2, 3, 11 and 12 as the bands that are sensitive to salinity. Let's see if we can use them as covariables and try to predict salinity using linear regression. 

```{r}
model1 <- lm(GCE_Salinity ~ ï..B1 + B2 + B3 + B11 + B12, data=salinity_data_filtered)
summary(model1)
```

Clearly the model is not a good fit. None of the variables are significant and the model R-sq is only 9.6%, meaning the chosen covariables only explain less than 10% of the variation in WSS. Individually, the covariables are explaining WSS by a very small magnitude (so we won't discuss inference of each coviariable separately for this model). We saw in the previous section that Bands 2 and 3 are highly correlated. Let's exclude Band 3 as a covariable and see if that model does any better. 

```{r}
model2 <- lm(GCE_Salinity ~ ï..B1 + B2 + B11 + B12, data=salinity_data_filtered)
summary(model2)
```

There is no significant change in the model performance except B2 becoming a significant covariable.

> Section Summary: We found out what bands are most sensitive to salinity variation. Then we tried to use those bands to model salinity, but the models were not a good fit. This indicates that we might need more covariables except the bands to model WSS.

### 2. Can we use literary proven remote sensing indices to model WSS?
Literature links WSS variation with changes in CDOM, phytoplankton blooms (NDCI), turbidity (NSMI) and three band ratios (B4/B3, B4/B2, B4/B1). Let's see if we can use them as covariables to estimate WSS. Before that, let's plot a correlation matrix for these variables to check for any high correlation among the independent variables.

```{r}
# filtering data points only to indices
salinity_indices <- salinity_data_filtered %>%
  select(CDOM.Proxy..B3.B4., NDCI, NSMI, B4.B3, B4.B2, B4.B1, GCE_Salinity)

# making sure all we have is indices and salinity
pander(head(salinity_indices))

# we know that all of these variables are not normal so again we will use spearman correlation 
CM <- cor(salinity_indices, method = "spearman")
corrplot(CM, method = "color", addCoef.col = "grey")
```

Even though a few covariables are explaining each other well, we will not discard any of them owing to the literature. Also, it appears that CDOM is simply inverse of B4/B3, so we can get rid of CDOM as a covariable. Let's see if we can estimate salinity using this set of covariables and a linear model.

```{r}
# removing cdom
salinity_indices <- salinity_indices %>%
  select(-CDOM.Proxy..B3.B4.)

# model
model3 <- lm(GCE_Salinity ~ NDCI + NSMI + B4.B3 + B4.B2 + B4.B1, data = salinity_indices)
summary(model3)
```

This model certainly did way better than the band-based models. We can observe that not only is the model significant, but the R-sq shoots up to 44%, which is pretty decent. Two of the chosen covariables are significant - NSMI and B4/B2. Discussing about how the covariables are explaining the variation in WSS:

* for every 1 unit increase in NDCI, WSS increases by 6 units, which is interesting to me as literature points out that dying phytoplankton affects WSS. 
* for every 1 unit increase in NSMI, WSS increases by 122 units, which in fact makes sense despite 122 being a big jump in salinity units. This is because NSMI is an index that ranges from 0-1, 1 being the most turbid. In real life, we wouldn't see an NSMI value change from 0 to 1 within a very small spatial area -- it *might* happen when we come from a central part of ocean to the marshy coast
* for every 1 unit increase in B4/B3, WSS increases by 3 units
* for every 1 unit increase in B4/B2, WSS decreases by 52 units. This is interesting to me and makes me wonder if the band ratio B4/B2 is a better proxy for CDOM instead of B4/B3, because there is consistent literature pointing out and using inverse relationship between WSS and salinity 
* for every 1 unit increase in B4/B1, WSS decreases by 1 unit. Using B1 is always tricky business as it containes the highest scattering signal from the atmosphere

Let's look at the model residuals and check the model for multicollinearity and heteroskedasticity 

```{r}
# normality of residuals
ggplot(model3, aes(sample=model3$residuals)) + stat_qq() + stat_qq_line() + ggtitle("WSS~Indices Residuals")

# multicollinearity 
vif(model3)


#heteroskedasticity 
bptest(model3)
```

As we can see here:

* The model residuals are not normal (although they are for the most part, they converge into a frown) and are negatively skewed, which indicates that the outliers are creating some bias in the model.
* There is definitely multicollinearity in the data, since the vif factor is not between 0-2 for all covariables. 
* Moreover, the bp-test also indicates that the data is heteroskedastic, since p-value  is significant and our hypothesis that the data is not heteroskedastic is rejected. 

So this is evidently not the best model for estimating WSS from satellite data, but seemingly not the worst either. And it does answer our question -- it is definitely possible to estimate WSS using the literary proven indices and band ratios. Let's see if we can make a better model by including the bands that we identified in the previous section.

```{r}
# filtering selected covariables
salinity_final <- salinity_data_filtered %>%
  select(ï..B1, B2, B3, B11, B12, NDCI, NSMI, B4.B3, B4.B2, B4.B1, GCE_Salinity)

# verifying
pander(head(salinity_final))

# model
model4 <- lm(GCE_Salinity ~ ï..B1 + B2 + B3 + B11 + B12 + NDCI + NSMI + B4.B3 + B4.B2 + B4.B1, data = salinity_final)
summary(model4)
```

This is definitely an improvement from the previous model that we had, as now the covariables explain more than 50% of the variation in salinity, which is not too bad an R-sq (0.52). However, only one of the covariables, B4.B2 is significant this time. Bands are not explaining the variation with a higher magnitude, however, by including all the bands as covariables, I feel that the indices are explaining the WSS variation way more realisitically (I will only point out specific covariables):

* For every 1 unit increase in NDCI, WSS decreases by ~2 units. This makes more sense to me now, as dying phytoplankton is supposed to affect WSS and here, as NDCI increases, phytoplankton is increasing and not dying off, so less turbidity and hence decreased salinity
* For every 1 unit in NSMI, WSS increases by 9 units. Again, a way more realistic increase in salinity due to turbidity as compared to the previous model

Let's run some model diagnostics:

```{r}
# normality of residuals
ggplot(model4, aes(sample=model4$residuals)) + stat_qq() + stat_qq_line() + ggtitle("WSS~Indices + Bands: Residuals")

# multicollinearity
vif(model4)

# heteroskedasticity
bptest(model4)
```

So evidently, there are still issues with the model:

* Residuals are not entirely normal 
* Definite multicollinearity and hetereoskedasticity issues in the data

While this model could still use some fine-tuning, it can serve as a base to build up more on how we can estimate WSS entirely from remote sensing data. 

> Section Summary: We tried to use indices to predict salinity and found that they do a way better job than bands, following which we tried to model salinity using both bands and indices which definitely improved the model.

## Discussion
Most of the inferences are linked in the previous section, so this section focuses on answering the research questions of the project.

1. *Is it possible to estimate water surface salinity by remote sensing methods using satellite data, specifically along the coast of Georgia? If yes, what wavelength is sensitive to WSS variation?*

Yes, it is definitely possible to estimate WSS by using entirely remote sensing data. To account for the fact that the salinity estimation is tuned for coastal regions, the salinity data was filtered for salinity values only greater than 2, getting rid of all freshwater. While WSS does not have any evident  Statistical analysis indicated that among all the Sentinel-2 bands, Band-2 is the most sensitive to salinity variation. Band-2 is at 494 nm, so we can say that far blue region of the optical spectrum is most sensitive to salinity variation.

2. *Moreover, since WSS depends highly on the turbidity and suspended particles in the water, can remote sensing based indices/quantities such as Color Dissolved Organic Matter (CDOM), Normalized Difference Chlorophyll Index (NDCI) and Normalized Suspended Material Index (NSMI) be used as proxy to estimate WSS from satellite data?*

Yes, the remote sensing based indices actually do a pretty decent job at predicting WSS. Statistical analysis indicated that a linear model comprising only of the indices as covariables explained upto 44% of the WSS variation. Moreover, it was discovered that the 'CDOM' variable was an inverse of the band ratio B4/B3. Since CDOM is known to have an inverse relationship with WSS, B4/B3 might have been an incorrect proxy for it and B4/B2 instead may be a better proxy.

The modelling for this project was carried out in steps and model diagnostics run for all models. It was discovered that by using a combination of both bands and indices as covariables to linear model WSS, more than 50% of WSS variation is being explained. This model, while still having issues of multicollinearity and  heteroskedasticity, can definitely serve as a starting point when it comes to predicting WSS from satellite-based reflectance data. All in all, this study in itself exolores the whole idea of estimating WSS from satellite data with some initial statistical analysis and concludes that even though WSS does not have an evident color signal, it is indeed possible to predict WSS from space! 

## Self-reflection
I really had a lot of fun doing this project mainly because a) I was able to make sure that I implement everything that I was hoping to learn out of this course in this project, giving me a guage of how far I was able to come b) This is also a research that I am working on and it was really exciting to see the initial results. This project spans across almost all the labs (maybe except the mapping ones) and coveres reading data, filtering data, data pre-analysis and assessing the normality (more importantly why), modelling univariate and multivariate regressions, understanding and inferring the models and also running model diagnostics to explore underlying issues because a high r-sq doesn't always mean a really good mode. I remember I asked you one question -- why use both qq-plots and histograms, and I think this project perfectly answered that question for me (Section 0 of Analysis). Moreover, I had a challenging time in Lab-7 in model inference and heteroskedasticity, so I made sure to have them as a requirement for this project -- it helped me understand them better and infer models correctly. Overall, I think I did a pretty good job on the project, and I think I'm a lot less intimidated now when it comes to statistics. THANK YOU SO MUCH for such a wonderful semester, Dr. Shannon! 